---
title: "Demographic Data"
author: "Matt Williamson"
date: "May 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r envPrep}
packages = c("tigris","tidycensus","raster","fasterize","tidyverse", "sf", "velox")

#use this function to check if each package is on the local machine
#if a package is installed, it will be loaded
#if any are not, the missing package(s) will be installed and loaded
package.check <- lapply(packages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

options(tigris_use_cache = TRUE)
census_api_key('baaaffb5ed3accd8dfa53c6f827659d43fcdfa21') #get this from the census api webpage see help(census_api_key) for details

prj <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"

outfolder <- "G:/My Drive/CONUSncedData/Census/"
rasterOptions(tmpdir = "D:/RastTemp/", todisk=TRUE)
```


# Tracts
## R Earnings functions
I developed a custom function that downloads the median income estimate from the 1990 and 2000 Decennial Census and the 2010 and 2016 American Community Survey. The value is converted to a raster based on the geometry of interest (here the tract) and adjusted for inflation (using the CPI inflation calculator https://www.bls.gov/data/inflation_calculator.htm). I then stack these rasters and extract their values based on current Census geography

```{r cars}
get_earnings_tct <- function(st) {
  #get state geometry for raster
  st.poly <- tigris::states() %>%  as(.,"sf") %>% filter(., STUSPS == st)  %>% st_transform(., crs= prj ) 
  st.rst <- fasterize::raster(st.poly, resolution=90)
  #get 1990 earnings data
  tct.1990 = tracts(state=st, year=1990, cb=TRUE)  %>% as(.,"sf") %>%  st_transform(., crs= prj ) %>% 
    mutate(.,GEOID = paste0(ST, CO, TRACTBASE,TRACTSUF))
  cty = as.vector(unique(tct.1990$COUNTYFP))
  e.1990 = reduce(
    map(cty, function(x) {
      tidycensus::get_decennial(geography = "tract", year=1990, variables="P080A001", state=st, county = x)
    }), rbind)
   tct.1990 <- tct.1990 %>% 
    geo_join(., e.1990, "GEOID", "GEOID")
  rst.1990 = fasterize(tct.1990, st.rst, field="value") 
  rst.1990.adj = round(rst.1990 * 1.8757, digits=2) #adjus for inflation base on Buerau of Labor Stats
  names(rst.1990.adj) = "earn1990"
  rm(rst.1990)
#get 2000 data
  tct.2000 = tracts(state = st, year = 2000, cb=FALSE) %>% as(.,"sf") %>%  st_transform(., crs= prj )
  e.2000 = tidycensus::get_decennial(geography = "tract", year= 2000, sumfile="sf3",variables = "P056001", state=st)
  tct.2000 <- tct.2000 %>%   
      geo_join(., e.2000, "CTIDFP00", "GEOID") 
    rst.2000 <- fasterize(tct.2000, st.rst, field="value")
    rst.2000.adj = round(rst.2000 * 1.4054, digits=2)
    names(rst.2000.adj) <- "earn2000"
    rm(rst.2000)
#get 2010 data
    e.2010 = tidycensus::get_acs(geography = "tract", year=2010, variables = "B19049_001", state=st) 
    tct.2010 = tracts(state = st, year = 2010, cb=FALSE) %>% 
      geo_join(., e.2010, "GEOID10", "GEOID") %>% as(.,"sf") %>%  st_transform(., crs= prj )
    rst.2010 <- fasterize(tct.2010, st.rst, field="estimate")
    rst.2010.adj = round(rst.2010 * 1.0953, digits=2)
    names(rst.2010.adj) = "earn2010"
    rm(rst.2010)
#get 2016 data
    e.2016 = tidycensus::get_acs(geography = "tract", year=2016, variables = "B19049_001", state=st) 
    tct.2016 = tracts(state = st, year = 2016, cb=FALSE) %>% 
      geo_join(., e.2016, "GEOID", "GEOID") %>% as(.,"sf") %>%  st_transform(., crs= prj )
    rst.2016 <- fasterize(tct.2016, st.rst, field="estimate")
    names(rst.2016) <- "earn2016"
 #stack and extract
    earn.stack = raster::stack(rst.1990.adj, rst.2000.adj, rst.2010.adj, rst.2016)
    rm(rst.1990.adj, rst.2000.adj, rst.2010.adj, rst.2016)
    gc()
    writeRaster(earn.stack, filename=paste0(outfolder,st,"earn.tif"), options="INTERLEAVE=BAND", overwrite=TRUE)
    vx.stack = velox(earn.stack)
    rm(earn.stack)
    gc()
    df = vx.stack$extract(sp=tct.2016, fun=NULL, df=TRUE)
    colnames(df) <- c("ID","earn1990","earn2000", "earn2010", "earn2016")
    id.lookup <- data.frame(ID=seq_along(tct.2016$GEOID),
                            GEOID = tct.2016$GEOID)
    df.lookup <- df %>% left_join(., id.lookup, by="ID") %>% 
      gather(., key=year, value=earnings, earn1990, earn2000, earn2010, earn2016) %>% 
      group_by(GEOID,year) %>% 
      summarise(., mnEarn=mean(earnings),
                sdEarn = sd(earnings),
                maxEarn = max(earnings),
                minEarn = min(earnings))
gc()
}

```

```{r}
state.omit <- c("AK", "HI", "DC", "AS", "GU", "MP", "PR", "UM", "VI","FL","MT")
conus <- fips_codes %>% filter(., !state %in% state.omit) %>% distinct(.$state) %>% c() %>% unlist()


tct.earn <- map(conus, function(x){
  get_earnings_tct(x)
})

